Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2015-10-23 latex/article.tex

\begin{figure}
  \begin{center}
    {\includegraphics[width=2.7cm]{figs/Octa_mean}}
    {\includegraphics[width=2.7cm]{figs/Octa_gaussian}}
    {\includegraphics[width=2.7cm]{figs/Octa_k1}}
    {\includegraphics[width=2.7cm]{figs/Octa_k2}}\\
    {\includegraphics[width=2.7cm]{figs/Octa_dir_min}}
    {\includegraphics[width=2.7cm]{figs/Octa_dir_max}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_min_zoom}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_max_zoom}}\\
    % {\includegraphics[height=4.0cm]{figs/EG_dragon_normal_zoom}}
    {\includegraphics[width=2.2cm]{figs/function_mean_0}}
    {\includegraphics[width=2.2cm]{figs/function_mean_1}}
    {\includegraphics[width=2.2cm]{figs/function_mean_2}}
    {\includegraphics[width=2.2cm]{figs/function_mean_3}}
    {\includegraphics[width=2.2cm]{figs/function_mean_4}}
  \end{center}
  \caption{\emph{Third row:} Mean, Gaussian, first and second principal curvature estimation on an OctaFlower with a digital domain of $130^3$.
  \emph{Second row:} First and second principal directions estimation (and zoom) on an OctaFlower.
  \emph{Third row:} Mean curvature computed in real-time on a dynamic object.}
  \label{fig:full}
\end{figure}

We first evaluate the curvature estimation on a full-resolution geometry
obtained by disabling the LoD criterion at the mesh generation step.
Figure~\ref{fig:full}-top and middle shows results of curvature tensor
estimation (mean, Gaussian, first and second principal curvatures, first and
second principal directions) on an OctaFlower at level $l_0$ (ground truth).
Figure~\ref{fig:full}-bottom shows mean curvature estimation in real-time on a
dynamic object. For all illustrations, we use directly normal vectors computed
by algorithm discussed in Section~\ref{sec:curv-tens-estim}.

@@

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000) 
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000) 
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000) 
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000) 
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000) 
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20, 
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000) 
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000) 
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \caption{Timings in milliseconds obtained while visualizing an adaptive
  triangulation on three objects -- OctaFlower (digital domain of $130^3$ ), Snow
  Microstructures ($233^3$) and XYZ-Dragon ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  }
  %
  \label{fig:timings}
\end{figure}

In this experiments, we measure the timings obtained when running our framework on a
GeForce GTX 850M. Here, the triangulation is dynamically refining according to its distance to the camera.
At each frame, both the curvature and the triangulation are recomputed.
If the visualisation stays unchanged, we compute the curvature at a level $l-1$ until the level $l_0$ is reached.
We measured those timings using the regular and the hierarchical algorithm. The results are presented in 
logarithmic scale figure \ref{fig:timings}.

We can first note that the required time for an object to compute ground truth curvature is usually
as high as the time required to extract the geometry and compute all the above levels.
Using approximations is thus mandatory, since for big objects such as the XYZ Dragon (Digital Domain $510^3$)
with high radii, the computation is seconds faster, the $l_0$ visualisation takes more than 4 seconds to compute, whereas
the approximated visualisation at $l_2$ takes only 300ms.
We can thus get an approximated visualisation of our object as soon as we run the application, allowing for
real time interactions with the object (changing the visualised value, the curvature radius, etc ...).
Once an interesting portion if the object is reached, it only takes a few seconds to the application to display
ground truth curvature estimation.

It is also clear on those graphs that using a hierarchical decomposition reduces the curvature computation time.
For the XYZ Dragon with a radius of 20, it needs almost 4 seconds to compute the $l_0$ level, whereas using the
hierarchical algorithm, it only needs 700ms. However, due to the precomputation, this algorithm is biased and
creates an error (presented in Table~\ref{tab:full-res-stat}) that needs to be considered

%\subsection{Visualisation}
%
%For dynamic datasets, we consider a time varying implicit
%surface. Since it would have been to expensive to update the a mipmap texture
%of densities at each time step, we simply evaluate the implicit
%expression at each vertex of a cell to decide if such cell is included
%or not into $\Ball{R}{\vx}\cap \Shape$.
%\todo[inline]{blabla experiments}

>>>>

\begin{figure}
  \begin{center}
    {\includegraphics[width=2.7cm]{figs/Octa_mean}}
    {\includegraphics[width=2.7cm]{figs/Octa_gaussian}}
    {\includegraphics[width=2.7cm]{figs/Octa_k1}}
    {\includegraphics[width=2.7cm]{figs/Octa_k2}}\\
    {\includegraphics[width=2.7cm]{figs/Octa_dir_min}}
    {\includegraphics[width=2.7cm]{figs/Octa_dir_max}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_min_zoom}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_max_zoom}}\\
    % {\includegraphics[height=4.0cm]{figs/EG_dragon_normal_zoom}}
    {\includegraphics[width=2.2cm]{figs/function_mean_0}}
    {\includegraphics[width=2.2cm]{figs/function_mean_1}}
    {\includegraphics[width=2.2cm]{figs/function_mean_2}}
    {\includegraphics[width=2.2cm]{figs/function_mean_3}}
    {\includegraphics[width=2.2cm]{figs/function_mean_4}}
  \end{center}
  \caption{\emph{First row:} Mean, Gaussian, first and second principal curvature estimation on an OctaFlower with a digital domain of $130^3$.
  \emph{Second row:} First and second principal directions estimation (and zoom) on an OctaFlower.
  \emph{Third row:} Mean curvature computed in real-time on a dynamic object.}
  \label{fig:full}
\end{figure}

We first evaluate the curvature estimation on a full-resolution geometry
obtained by disabling the LoD criterion at the mesh generation step.
Figure~\ref{fig:full}-top and middle shows results of curvature tensor
estimation (mean, Gaussian, first and second principal curvatures, first and
second principal directions) on an OctaFlower at level $l_0$ (ground truth).
Figure~\ref{fig:full}-bottom shows mean curvature estimation in real-time on a
dynamic object. For this case, we simply evaluate the implicit expression at
each vertex of a cell to decide if such cell is included or not into
$\Ball{R}{\vx}\cap \Shape$ instead of updating the \emph{mipmap texture} of
densities at each time step. For all illustrations, we use directly normal
vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}.

@@

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000)
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000)
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000)
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000)
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000)
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000)
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000)
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \caption{Timings in milliseconds (in logscale) obtained while visualizing an adaptive
  triangulation on three objects -- OctaFlower (digital domain of $130^3$ ), Snow
  Microstructures ($233^3$) and XYZ-Dragon ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  Timings are given using a NVIDIA GeForce GTX 850M GPU.
  }
  %
  \label{fig:timings}
\end{figure}

When dealing with large datasets, we can not expect real-time curvature tensor
estimation due to the huge amount of data to process. By computing an
dynamically refined approximate curvature tensor estimation joined with an
adaptive triangulation (as discussed in Section~\ref{sec:hierarchical}), we
manage to maintain a real time framerate by giving control over the amount of
data to process at each frame. In Figure~\ref{fig:timings}, we compare timings
(in logarithmic scale) when the triangulation is dynamically refining according
to its distance to the camera. We measured those timings using the regular and
the hierarchical algorithm.

We can first note that the required time for an object to compute ground truth
curvature is usually as high as the time required to extract the geometry and
compute all the above levels. Using approximations is thus mandatory, since for
big objects such as the XYZ-Dragon ($510^3$) with big radii the approximate
computation is faster than the $l_0$ while results are enough close to the
$l_0$. For example, with $R=20$, the approximated visualization at $l_2$ takes
only 300ms while the $l_0$ visualization takes more than 4 seconds.

We can thus get an approximated visualisation of our object as soon as we run
the application, allowing for real time interactions with the object (changing
the visualized value, curvature radius, \ldots). It is also clear on those
graphs that using a hierarchical decomposition reduces the curvature computation
time, especially with big radii. For the XYZ-Dragon with $R=20$, it needs almost
4 seconds to compute the $l_0$ level, whereas using the hierarchical algorithm,
it only needs 700ms. However, due to the precomputation and current hierarchical
structure, this algorithm is biased and creates an error (presented in
Table~\ref{tab:full-res-stat}) that needs to be considered.
